---
output:
  pdf_document: default
  word_document: default
date: "2024-10-12"
author: "Jane Lee"
---


## R Markdown


# Question 1a
y <- c(2, 5, 4.3, 4.2, 8) # Observed data
compute.mse <- Vectorize(function(b0) {
  return(mean((y - b0) ** 2))
})
b0 <- seq(0, 10, length.out = 10000 + 1) # Try various b0's
mse <- compute.mse(b0)

## Plot b0 vs mse
plot(b0, mse, type="l", col="blue", xlab="b0", ylab="MSE")

## (Q1a continued) Min MSE and corresponding b0
min_mse_index <- which.min(mse)
b0_optimal <- b0[min_mse_index]
mse_min <- mse[min_mse_index]

mse_min
b0_optimal

### (Q1a continued) b0 = mean(y)
y_mean = mean(y)
if (mean(y) == b0_optimal) {
  print("optimal b0 value equals to mean(y)")
}else {
  print("optimal b0 value does not equal to mean(y)")
}


```{r Q1b}
# Question 1b

y <- c(2, 5, 4.3, 4.2, 8) 

# Fit a linear model with only the intercept
model <- lm(y ~ 1)

# Summary of the model
summary(model)

```

```{r Q2a}
# Question 2a

y <- c(2, 5, 4.3, 4.2, 8) # Observed data
compute.mae <- Vectorize(function(b0) {
return(mean(abs(y - b0)))
})
b0 <- seq(0, 10, length.out = 10000 + 1) # Try various b0's
mae <- compute.mae(b0)

## Plot b0 vs mae
plot(b0, mae, type = "l", col = "red", xlab = "b0", ylab = "MAE") # Plot b0 vs mae

## (Q2a continued) The shape
print("The shape of the plot b0 vs. MAE is V-shaped. In between the y-values, the MAE gradually increases linearly and there are sharp transitions at the data points.")

```


```{r Q2b}
# Question 2b

## Min MAE and corresponding b0
min_mae_index <- which.min(mae)
b0_optimal1 <- b0[min_mae_index]
mae_min <- mae[min_mae_index]

mae_min
b0_optimal1

### (Q2b continued) b0 = median(y)
y_median <- median(y)

if (y_median == b0_optimal1) {
  print("optimal b0 value equals to median(y)")
}else {
  print("optimal b0 value does not equal to median(y)")
}
```


```{r Q3a}
# Question 3a

##Install packages
install.packages("ggplot2")
library(ggplot2)

##Read & clean csv data
chicago_taxi_raw <- read.csv("chicago_taxi_trip_sample.csv")
chicago_taxi <- na.omit(chicago_taxi_raw[c("trip_total", "trip_miles")])

##Linear regression model
model <- lm(trip_total ~ trip_miles, data = chicago_taxi)
cat("The linear regression model is: y =",coef(model)[1],"+",coef(model)[2],"x", "\n")

##MSE
y_pred <- predict(model, chicago_taxi)
mse <- mean((chicago_taxi$trip_total - y_pred)^2)
cat("Mean Squared Error (MSE):", mse, "\n")

##R-Squared
summary(model)
r_squared <- summary(model)$r.squared
cat("R-squared:", r_squared, "\n")

##Corr(Y,Y predicted)
y_obs <- chicago_taxi$trip_total
corr <- cor(y_obs, y_pred)
squared_corr <- corr^2
cat("Corr(Y,Y predicted) squared value:", squared_corr, "\n")

##Checking
if (squared_corr == r_squared) {
  print("Corr(Y,Y predicted) squared value equals to R-squared")
}else {
  print("Corr(Y,Y predicted) squared value does not equal to R-squared")
}

##Predicted Cost for 10 miles trip
predicted_cost_10_miles <- predict(model, data.frame(trip_miles = 10))
cat("\n","Predicted cost for a 10-mile trip:", predicted_cost_10_miles)

```

```{r Q3b}
# Question 3b

## Converting to Trip km
chicago_taxi$trip_km <- chicago_taxi$trip_miles * 1.609
model_km <- lm(trip_total ~ trip_km, data = chicago_taxi)

## B0 & B1
cat("Intercept (β0) for kilometers model:", coef(model_km)[1], "\n")
cat("Slope (β1) for kilometers model:", coef(model_km)[2], "\n")
cat("B0 (intercept) remains the same as the B0 value in Question 3a. However, B1 (slope) gets scaled by the conversion factor (1.609),with the slope in km being smaller than the slope in miles.", "\n")

### Predicted cost for 16.09 km trip
predicted_cost_16_km <- predict(model_km, data.frame(trip_km = 16.09))
cat("Predicted cost for a 16.09 km trip:", predicted_cost_16_km, "\n")
cat("This prediction matches the prediction for a 10-mile trip in Question 3a because 16.09 km is 10 miles.")

```

```{r Q3c}
# Question 3c

# Clean the variables
chicago_taxi_clean <- na.omit(chicago_taxi_raw[c("trip_total", "trip_miles", "tolls", "extras")])

## Model with trip_miles & tolls: R-squared
lmodel_miles_tolls <- lm(trip_total ~ trip_miles + tolls, data = chicago_taxi_clean)
r_squared_miles_tolls <- summary(lmodel_miles_tolls)$r.squared
cat("R-squared for trip_miles + tolls:", r_squared_miles_tolls, "\n")

## Model with trip_miles & extras: R-squared
lmodel_miles_extras <- lm(trip_total ~ trip_miles + extras, data = chicago_taxi_clean)
r_squared_miles_extras <- summary(lmodel_miles_extras)$r.squared
cat("R-squared for trip_miles + extras:", r_squared_miles_extras, "\n")

## More useful variable: tolls vs. extras
useful <- if (r_squared_miles_tolls > r_squared_miles_extras) {
  print("More useful explanatory variable based on R-squared: tolls")
} else {
  print("More useful explanatory variable based on R-squared: extras")
}

## Model with trip_miles & tolls & extras: R-squared
model_miles_tolls_extras <- lm(trip_total ~ trip_miles + tolls + extras, data = chicago_taxi_clean)
r_squared_miles_tolls_extras <- summary(model_miles_tolls_extras)$r.squared
cat("\n", "R-squared for trip_miles + tolls + extras:", r_squared_miles_tolls_extras, "\n")
cat("The R-Squared value is higher than the previous models. This higher R-squared value suggests that adding tolls and extras, along with trip_miles, improves the model's ability to explain the variation in total trip cost.")

```
